---
title: "RI Wood Turtle Data"
output: html_notebook
---

# Introduction
The goal of this project is to analyze telemetry data collected over to 2022 field season on six wood turtles (3 males and 3 females), to look for patterns and differences between their homeranges and how far away the areas they are occupying are from the stream.

## Getting started
Activate necessary packages and clear R's brain
```{r}
rm(list = ls())
library(dplyr)
library(ggplot2)
library(here)
library(ggfortify)
library(adehabitatHR)
library(sp)
```

## Pulling in the data
The telemetry data is the first dataset that needs to be pulled in and checked for errors

```{r}
turtles <- read.csv(here("Data", "RI_turtle_data.csv"), stringsAsFactors = TRUE)
```

## Check out the data

```{r}
str(turtles) #shows the data frame
```

Check the column names

```{r}
names(turtles)
```

Check the levels for the columns Turtle_ID and Turtle_sex. There should be six and two levels respectively.

```{r}
levels(turtles$Turtle_ID)
levels(turtles$Turtle_sex)
```

The Latitudes and Longitudes can be checked for errors by using histograms

First, the latitudes

```{r}
ggplot(turtles, aes(x = Latitude)) +
  geom_histogram(bins = 10)
```
Looks good! Slight variation is expected and the large number of counts in the center is great as well because these turtles all live along the same river.

Now, the longitudes

```{r}
ggplot(turtles, aes(x = Longitude)) +
  geom_histogram(bins = 10)
```
Looks great!

## Calculating Homerange

Before we can calculate the home ranges for each turtle, we need to generate subsets that sort the data by each individual.

```{r}
L1R2_male <- filter(turtles, Turtle_ID == "L1R2")
L3R1_male <- filter(turtles, Turtle_ID == "L3R1")
L9R0_male <- filter(turtles, Turtle_ID == "L9R0")
L1R4_female <- filter(turtles, Turtle_ID == "L1R4")
L3R3_female <- filter(turtles, Turtle_ID == "L3R3")
L3R9_female <- filter(turtles, Turtle_ID == "L3R9")
```

In order for R to interpret the latitudes and longitudes as coordinates, a matrix needs to be created with the X and Y values for the coordinate locations. Lets work through the entire workflow with L1R2 first

```{r}
L1R2_latlong <- cbind(L1R2_male$Latitude, L1R2_male$Longitude)
```

Now convert the matrix into spatial points

```{r}
L1R2_sp <- SpatialPoints(L1R2_latlong)
L1R2_sp
```

Now that R recognizes the coordinates for L1R2, we need to use the clusthr() function to implement a single-linkage clustering algorithm 

```{r}
L1R2_clu <- clusthr(L1R2_sp)
class(L1R2_clu)
```

Plot the results
```{r}
plot(L1R2_clu)
```
Now put together the MCP for L1R2

```{r}
L1R2_mcp <- mcp(L1R2_sp, percent = 95)
class(L1R2_mcp)
```

Plot the MCP for L1R2
```{r}
plot(L1R2_mcp)
plot(L1R2_sp, add=TRUE)
```

Now that we know the coding process for generating an MCP, replicate the process with the other individuals. 

Create the latlong groupings for the remaining 5 individuals

```{r}
L1R4_latlong <- cbind(L1R4_female$Latitude, L1R4_female$Longitude)
L3R1_latlong <- cbind(L3R1_male$Latitude, L3R1_male$Longitude)
L3R3_latlong <- cbind(L3R3_female$Latitude, L3R3_female$Longitude)
L3R9_latlong <- cbind(L3R9_female$Latitude, L3R9_female$Longitude)
L9R0_latlong <- cbind(L9R0_male$Latitude, L9R0_male$Longitude)

```

Convert into spatial points

```{r}
L1R4_sp <- SpatialPoints(L1R4_latlong)
L3R1_sp <- SpatialPoints(L3R1_latlong)
L3R3_sp <- SpatialPoints(L3R3_latlong)
L3R9_sp <- SpatialPoints(L3R9_latlong)
L9R0_sp <- SpatialPoints(L9R0_latlong)
```

Use the clusthr() function to implement a single-linkage clustering algorithm

```{r}
L1R4_clu <- clusthr(L1R4_sp)
L3R1_clu <- clusthr(L3R1_sp)
L3R3_clu <- clusthr(L3R3_sp)
L3R9_clu <- clusthr(L3R9_sp)
L9R0_clu <- clusthr(L9R0_sp)

```

Let's plot each of these

L1R4:
```{r}
plot(L1R4_clu)
```

L3R1:
```{r}
plot(L3R1_clu)
```

L3R3:
```{r}
plot(L3R3_clu)
```

L3R9:
```{r}
plot(L3R9_clu)
```

L9R0:
```{r}
plot(L9R0_clu)
```

Create the MCP for each of the remaining individuals
```{r}
L1R4_mcp <- mcp(L1R4_sp, percent = 95)
L3R1_mcp <- mcp(L3R1_sp, percent = 95)
L3R3_mcp <- mcp(L3R3_sp, percent = 95)
L3R9_mcp <- mcp(L3R9_sp, percent = 95)
L9R0_mcp <- mcp(L9R0_sp, percent = 95)
```

Plot the MCP for each individual

L1R4:
```{r}
plot(L1R4_mcp)
plot(L1R4_sp, add=TRUE)
```

L3R1:
```{r}
plot(L3R1_mcp)
plot(L3R1_sp, add=TRUE)
```

L3R3:
```{r}
plot(L3R3_mcp)
plot(L3R3_sp, add=TRUE)
```

L3R9:
```{r}
plot(L3R9_mcp)
plot(L3R9_sp, add=TRUE)
```

L9R0:
```{r}
plot(L9R0_mcp)
plot(L9R0_sp, add=TRUE)
```

Check out the area of the MCPs

L1R2:
```{r}
L1R2_area <- mcp.area(L1R2_sp, percent = seq(75,100, by = 5),
                      unin = c("km"),
                      unout = c("m2"), plotit = TRUE) #creates visualization of home range expansion 75% to 100% in increments of 5%
          

L1R2_area #examines dataframe
```

L1R4:
```{r}
L1R4_area <- mcp.area(L1R4_sp, percent = seq(75,100, by = 5),
          unin = c("km"),
         unout = c("m2"), plotit = TRUE)
L1R4_area
```

L3R1:
```{r}
L3R1_area <- mcp.area(L3R1_sp, percent = seq(75,100, by = 5),
          unin = c("km"),
         unout = c("m2"), plotit = TRUE)
L3R1_area
```

L3R3:
```{r}
L3R3_area <- mcp.area(L3R3_sp, percent = seq(75,100, by = 5),
          unin = c("km"),
         unout = c("m2"), plotit = TRUE)
L3R3_area
```

L3R9:
```{r}
L3R9_area <- mcp.area(L3R9_sp, percent = seq(75,100, by = 5),
          unin = c("km"),
         unout = c("m2"), plotit = TRUE)
L3R9_area
```

L9R0:
```{r}
L9R0_area <- mcp.area(L9R0_sp, percent = seq(75,100, by = 5),
          unin = c("km"),
         unout = c("m2"), plotit = TRUE)
L9R0_area
```

Merge areas into one dataset

What are the units of area? It is important to know.

```{r}
areas <- cbind(L1R2_area, L1R4_area, L3R1_area, L3R3_area, L3R9_area, L9R0_area)
colnames(areas) <- c("L1R2", "L1R4", "L3R1", "L3R3", "L3R9", "L9R0") 
areas
```
Ok! This is great - you've got different percent MCP home range areas for all 6 turtles in a data frame. You could pivot longer if you wanted to compare mean 95% male vs. female home ranges, for example. What kind of statistical test would that be?

I'm really impressed at how you moved through all this data and figured out how to spatilaize it. I'm not sure where we are with regard to looking at mean distance to river, or comparing mean dist to river b/w males and females.  What are the statistical tests you'd like to run here?


