---
title: "Distance_to_stream.Rmd"
author: "Julia Sirois"
date: "11/19/2022"
output: html_document
---

# Introduction
This .Rmd will work through code to compare the relationship between the distance to the Wood River and Flat River being occupied by males vs females. The distances used in this dataset were calculated using ArcGIS Pro

## Getting Started
```{r}
rm(list = ls())
library(here)
library(tidyverse)
library(ggfortify)
```

## Pulling in data
```{r}
turtles <- read.csv(here("Data", "updated_turtle_locations.csv"))
```

Create subsets for each sex in order to explore the data

Check column names
```{r}
names(turtles)
```

Create subset
```{r}
df <- dplyr::select(turtles, Turtle_sex, NEAR_DIST)
```


Plot the relationship
```{r}
ggplot(df, aes(NEAR_DIST))+
  geom_histogram() +
  facet_wrap(~Turtle_sex, ncol = 1) +
  theme_bw()
```
When you look at this, would you guess there is a difference? What would you guess the means to be for each sex?


Summarise the data
```{r}
turtlemeans <- summarise(
  group_by(df, Turtle_sex),
  meanNearDist = mean(NEAR_DIST))
```
You might want to add sd to that summary

Run ttest
```{r}
turtle_ttest <- t.test(NEAR_DIST ~ Turtle_sex, data = df)
turtle_ttest
```
There is a significant difference between the distance from the stream occupied by males vs females (n=6)(p=0.00)

Good. So now you need a "publication quality" plot that highlights the statistical difference for the reader. What might that look like?
